---
title: "causalCalibration Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{causalCalibration Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
if (!requireNamespace("causalCalibration", quietly = TRUE)) {
  repo_root <- normalizePath(file.path(getwd()), winslash = "/", mustWork = TRUE)
  if (!dir.exists(file.path(repo_root, "R")) && dir.exists(file.path(repo_root, "..", "..", "..", "..", "R"))) {
    repo_root <- normalizePath(file.path(repo_root, "..", "..", "..", ".."), winslash = "/", mustWork = TRUE)
  }
  source(file.path(repo_root, "r", "causalCalibration", "R", "utils.R"))
  source(file.path(repo_root, "r", "causalCalibration", "R", "methods.R"))
  source(file.path(repo_root, "r", "causalCalibration", "R", "fit_calibrator.R"))
  source(file.path(repo_root, "r", "causalCalibration", "R", "fit_cross_calibrator.R"))
  source(file.path(repo_root, "r", "causalCalibration", "R", "diagnostics.R"))
} else {
  library(causalCalibration)
}
set.seed(11)
```

# Overview

This vignette shows:

- simulate a paper-style treatment-effect problem,
- construct raw and cross-fitted prediction objects,
- fit ordinary calibration,
- fit cross-calibration in sample,
- inspect overlap before choosing the calibration target,
- compare predictions,
- and interpret diagnostics.

# Simulate a paper-style problem

```{r}
n <- 180
w <- runif(n, -1, 1)
propensity <- plogis(1.6 * w)
treatment <- rbinom(n, size = 1, prob = propensity)

tau_true <- 0.4 + 0.9 * w + 0.25 * sin(3 * w)
mu0 <- -0.2 + 0.6 * w
mu1 <- mu0 + tau_true
outcome_mean <- (1 - propensity) * mu0 + propensity * mu1
outcome <- ifelse(treatment == 1, mu1, mu0) + rnorm(n, sd = 0.18)

tau_raw <- 0.55 + 1.35 * tau_true + 0.25 * (w > 0) - 0.15 * cos(4 * w)

k_folds <- 5
fold_ids <- ((seq_len(n) - 1) %% k_folds) + 1
tau_fold_matrix <- sapply(seq_len(k_folds), function(fold) {
  tau_raw + 0.08 * (fold - 3) + 0.05 * sin((fold + 1) * w)
})
tau_oof <- tau_fold_matrix[cbind(seq_len(n), fold_ids)]
tau_new <- tau_raw + 0.03 * sin(6 * w)
```

# Ordinary calibration

```{r}
summary(assess_overlap(treatment = treatment, propensity = propensity))

calibrator <- fit_calibrator(
  predictions = tau_raw,
  treatment = treatment,
  outcome = outcome,
  mu0 = mu0,
  mu1 = mu1,
  propensity = propensity,
  loss = "dr",
  method = "isotonic"
)

tau_calibrated <- predict(calibrator, tau_new)
summary(calibrator)
head(tau_calibrated)
```

# Cross-calibration

- `predictions = tau_oof` are the pooled out-of-fold predictions used to fit the calibration map.
- `fold_predictions = tau_fold_matrix` is the full `n x K` matrix used for prediction-time aggregation.
- `fold_ids` identifies which column is the OOF prediction for each observation and lets the package validate the cross-fit bundle.

```{r}
validate_crossfit_bundle(
  predictions = tau_oof,
  fold_predictions = tau_fold_matrix,
  fold_ids = fold_ids
)

cross_calibrator <- fit_cross_calibrator(
  predictions = tau_oof,
  fold_predictions = tau_fold_matrix,
  fold_ids = fold_ids,
  treatment = treatment,
  outcome = outcome,
  mu0 = mu0,
  mu1 = mu1,
  propensity = propensity,
  loss = "dr",
  method = "isotonic"
)

tau_cross_calibrated <- predict(cross_calibrator, tau_fold_matrix)
summary(cross_calibrator)
head(tau_cross_calibrated)
```

# Compare predictions

```{r}
head(data.frame(
  raw = round(tau_raw, 3),
  ordinary = round(predict(calibrator, tau_raw), 3),
  cross = round(tau_cross_calibrated, 3),
  truth = round(tau_true, 3)
))
```

# Diagnostics

```{r}
diagnostics <- diagnose_calibration(
  predictions = tau_cross_calibrated,
  comparison_predictions = tau_raw,
  treatment = treatment,
  outcome = outcome,
  mu0 = mu0,
  mu1 = mu1,
  propensity = propensity,
  curve_method = "histogram",
  fold_ids = fold_ids,
  target_population = "both"
)

summary(diagnostics)
```

# What population is being calibrated?

The example above uses `loss = "dr"`, which targets calibration in the original study population. It relies on inverse-propensity weighting and can become less stable under weak overlap.

If overlap is weak, `loss = "r"` may be more robust because the R-loss weighting emphasizes the overlap region. The tradeoff is interpretational: the calibrated scores then target an overlap-weighted population rather than the original study population.

- choose `dr` when original-population targeting matters and overlap is adequate,
- choose `r` when overlap is weak and overlap-weighted targeting is acceptable.
